The Role of Video Data Sharing in Advancing Education Policy and Practice

### 2015-05-08 2015-06-01-nsf-ed-workshop/paper.tex
fix authors
1st pass thru Technical Challenges

\title{The Role of Video Data Sharing in Advancing Education Policy and Practice}
\shorttitle{Video Data Sharing in Education}
\author{Rick O. Gilmore}
\author{Karen E. Adolph}
\author{David S. Millman}
\affiliation{The Databrary (databrary.org) Project}

@@

Video is a uniquely rich medium for capturing in real time the interactions that occur in classrooms, muesuems, laboratory, home, and other informal learning settings.
Many researchers and practitioners in the developmental, learning, and education sciences collect video as raw data.
Video recordings capture what teachers say and do and what students say and do in ways that offer the potential to discover when, why, and how learning actually works.
Still, the open sharing of video data is rare, even within labs or among collaborators.
Video data should be conducive to sharing.
Video captures behavior that others can readily inspect.
It captures multiple dimensions of complex real-world learning settings in ways that others can easily build upon. But, technical, ethical, and cultural barriers have made open sharing of video data rare.
Videos come in varied formats and generage large files; these pose conversion and storage challenges.
Personally identifying information can be removed from text-based data.
Videos may contain faces, voices, names spoken aloud, views or voices of non-participants, views of classroom interiors and sometimes views of the homes of research participants. 
These elements cannot be removed without reducing the information content and reuse value to others. 
The collection of video or other identifiable or sensitive information requires approval by a research ethics board.
It also requires informed consent or assent from the participants, and possibly parental, teacher, and school district approval.
Researchers risk violating participants’ privacy if digital images are viewed or released to others without authorization.
For these and other reasons, no culture of widespread open sharing of video data has emerged in the education, learning, and developmental sciences.

@@

\section{The Promise of Video Data Sharing}

More than 1 billion users worldwide uploade more than 300 hours of video to YouTube every minute (https://www.youtube.com/yt/press/statistics.html). Google, YouTube's owner, has built an immense data infrastructure to upload, store, convert, tag, and stream video. The current scale of video collection in developmental and education contexts is undoubtedly smaller, but it is not small. The Gates Foundation-funded Measures of Effective Teaching (MET) Project, hosted at ICPSR, contains X video segments from X classroom settings, representing tens of terabytes of data. In just over a year of operation, the Databrary digital library has collected more than 7,000 individual videos, representing 2,400 hours of recording, featuring more than 1,800 infant, child, and adult participants. Video data is big, and the interest in recording and sharing video for research, education, and policy purposes has grown.

\subsection{Video is Uniquely Rich}

\subsection{Video Uses in Education}

We record teachers in training and teachers in-service for evaluation.
We record classrooms to capture what teachers do and how students respond.
We record visitors and docents in museum settings.
\section{The Challenges of Video Data Sharing}

Sharing video poses challenges. 
These include technical infrastructure, research ethics, privacy, and security, and data management. 
Further, making open data sharing in education research an established norm requires changing understanding about how identifiable data might be shared and building a community of researchers and educators committed to making video data sharing work.

\subsection{Technical}

@@

\subsubsection{Video Data is Big}

Video often accompanies other data streams (physiological recordings, brain imaging, EMG, motion tracking, eye tracking, verbal transcripts). 
Multiple data streams—including two or more camera views—require tools for synchronization and integration and benefit from multivariate time series analyses. 
Many labs lack the resources to adopt tools for detailed video coding and complex analyses. 
Furthermore, widespread use of video creates a data explosion: The typical developmental lab collects 8-12 hours of video/week 18 in widely varied formats. 
Thus, sharing digital video requires substantial storage capacity, powerful search and streaming tools, and significant computational resources for transcoding videos into common, preservable formats.

\subsubsection{Tools for Coding Video}
After videos are recorded, human coders evaluate videos and apply text-based annotations or numeric codes to videos.
Coders use a wide-range of commercial and academic software tools: Transana, StudioCode, V-code, MaxQDA, Noldus Observer, Mangold Interact, and Datavyu, among others.
Some use spreadsheets or pencil and paper.
The tools have no standard format for data storage or export.
Some offer web/cloud based storage (e.g., Transana, V-code), but otherwise coded data may not be easily shared within tools, much less between them.
With few exceptions, machine learning has not been applied to video data coding.
\subsection{Ethics and Privacy}
Sharing video recordings poses a unique challenge to existing data sharing policies because videos contain personally identifying information—specifically participants’ faces and voices and often the
insides of their homes and classrooms. 
Sharing personally identifiable information puts research participants and others depicted in recordings at increased risk for loss of privacy. 
At the same time, blurring or altering original recordings to hide identities undermines or eliminates
their value to other researchers. 
Often, participants’ faces and voices produce the behaviors of interest.
A challenge for future education research is to find ways to preserve the value of video for reuse by others while protecting the privacy of research participants.
Databrary has developed an access model described below that provides an example of how this might be accomplished.
\subsection{Data Management}
A primary impediment to sharing is researchers’ lack of time to find, label, clean, organize, and copy their files into formats that can be used and understood by others15. 
Even scientists committed to data sharing find it difficult to do. Data management systems that make sharing convenient, reliable—and optimally automatic29—can help.

Existing data management practices pose challenges for video sharing.
Despite similar research methods, study designs vary widely. 
No two developmental science labs manage data in the same way. 
Some studies are longitudinal, where researchers observe the same participants at
multiple sessions. 
Some are cross-sectional, where researchers observe
each participant at only one session. 
The timing of observations may be
determined by participants’ age (e.g., 4-month-olds, 8-month-olds,
12-month-olds) or abilities (e.g., preverbal, one-word utterances,
sentences), experimentally determined variables (e.g., pre- and
post-intervention), or other factors. 
Some labs organize longitudinal data by grouping files first by participant and then by session date.
Other labs organize longitudinal data by session and then by
participant. 
Still others organize longitudinal data based on task (book reading, block building, free play). Some researchers institute a central data management system to be followed by all the lab members, providing easier access and greater transparency for the entire lab but not necessarily providing the structure for similar benefits to
researchers unfamiliar with the lab’s practices. 
Other researchers allow their students to keep separate records, making it difficult to share
data even within the lab. 
Some researchers keep videos, metadata, and analyses together, and some do not. Idiosyncratic terms, record-keeping, and data management systems are the norm. 
Databrary must enable researchers to discover and understand each other’s materials, regardless of the original investigator’s data management system.
As in other fields22, developmental science is inundated by an explosion of digital data, most of which is inaccessible to other researchers. The data deluge stems from the advent of cheap, high-resolution digital cameras. The average developmental lab collects 12 hours of video/week23. Raw data without information about workflow and data provenance is essentially “orphaned” and cannot be understood by subsequent researchers who weren’t part of the original team24. Moreover, video files must be electronically linked with participant permissions and relevant metadata. An effective data management system for video does not currently exist. We will build one to allow researchers within labs to benefit from each other’s work, facilitate interdisciplinary collaborations, and provide a solid foundation for open sharing25

@@

\section{Databrary.org}
Databrary (databrary.org) is a digital data library specialized for storing and sharing video and research data and metadata.
Databrary is a joint project of New York University (NYU) and The Pennsylvania State University (Penn State). 
The Databrary project began with a workshop on open video data sharing funded by the National
Science Foundation (NSF) under Grant No. BCS-1139702 at which developmental scientists, computer scientists, library scientists, and federal agency program officers discussed the promises and challenges of
sharing video data. 
NSF and the National Institute of Child Health and Human Development (NICHD) have provided funding under grant No. BCS-1238599 and Cooperative Agreement U01-HD-076595, respectively. 
The interdisciplinary team bring expertise in developmental and neural
science, information technology, library science, software development,
data curation, and community engagement. 
In addition, a board of expert
advisors composed of developmental scientists, library scientists, and
leaders of data repositories in the behavioral and social sciences
provide guidance.
No large-scale repository for sharing video data currently exists. So, we will create a web-based Databrary repository. 

@@

Databary is built on a PostreSQL database using the Scala Play framework and JavaScript.
Data are preserved indefinitely in a secure data storage facility at NYU managed by central IT.
There is no cost to use the system; an institutional subscription model is under development.
All code is hosted on GitHub.
Databrary can house video, audio, PDF, spreadsheet, image, and text-based files (along with associated metadata), as well as executable scripts in Ruby, R and Matlab that facilitate data analysis.
Video and audio data are transcoded into standard and HTML5-compatible formats, currently H.264+AAC as MP4. Databrary stores other data in native formats (e.g., .doc, .docx, .xls, .xlsx, .txt, .csv, .pdf, .jpg, .png).

>>>>

\title{The Role of Video Data Sharing in Advancing Education Policy and Practice}
\shorttitle{Video Data Sharing in Education}
\author{Rick O. Gilmore, Karen E. Adolph, David S. Millman}
\affiliation{The Databrary (databrary.org) Project}

@@

Video is a uniquely rich medium for capturing in real time the interactions that occur in classrooms, muesuems, laboratory, home, and other informal learning settings.
Many researchers and practitioners in the developmental, learning, and education sciences collect video as raw data.
Video recordings capture what teachers, docents, parents, tutors and peers say and do and what learners say and do in response.
The recordings capture these dynamics in ways that offer the potential to discover who did what, and when, why, and how it influenced learning.
Still, the open sharing of video data is rare, even within labs or among collaborators.
Video data should be conducive to sharing.
Video captures behavior that others can readily inspect.
It captures multiple dimensions of complex real-world learning settings in ways that others can easily build upon. But, technical, ethical, and cultural barriers have made open sharing of video data rare.
Videos come in varied formats and generage large files; these pose conversion and storage challenges.
Personally identifying information can be removed from text-based data.
Videos may contain faces, voices, names spoken aloud, views or voices of non-participants, views of classroom interiors and sometimes views of the homes of research participants. 
These elements cannot be removed without reducing the information content and reuse value to others. 
The collection of video or other identifiable or sensitive information requires approval by a research ethics board.
It also requires informed consent or assent from the participants, and possibly parental, teacher, and school district approval.
Researchers risk violating participants’ privacy if digital images are viewed or released to others without authorization.
For these and other reasons, no culture of widespread open sharing of video data has emerged in the education, learning, and developmental sciences.

@@

\section{The Promise of Video}

More than 1 billion users worldwide uploade more than 300 hours of video to YouTube every minute (https://www.youtube.com/yt/press/statistics.html). 
Google, YouTube's owner, has built an immense data infrastructure to upload, store, convert, tag, and stream video. 
The current scale of video collection in developmental and education contexts is undoubtedly smaller and more focused in purpose, but it is large and growing. 
The Gates Foundation-funded Measures of Effective Teaching (MET; http://www.metproject.org/) Project generated more than 1,000 full-length video segments from K-12 classrooms, covering core subjects such as mathematics, language arts from multiple camera angles. 
The data constituting tens of terabytes of storage are hosted at the University of Michigan (http://soe.mivideo.it.umich.edu/) and securely streamed to registered viewers across the country.
Similarly, in just over a year of operation, the Databrary (http://databrary.org) digital library has collected more than 7,000 individual videos, representing 2,400 hours of recording, and featuring more than 1,800 infant, child, and adult participants.
Databrary has more than 100 authorized researchers representing more than 50 institutions across the globe who access the videos.
Video data is big, and the interest in recording and sharing video for research, education, and policy purposes continues to grow.

Video captures the richness and complexity of human behavior and interactions between people and their environment like no other measure.
Video captures what people say and what they do.
It captures when and where they look or gesture, when, where and how they move, and how their looking, gesturing, and moving corresponds to what others are doing.
These characteristics mean that video has multiple use cases in the study of learning and development.
In laboratory contexts, researchers use videos to study how infants, children, and adults behave spontaneously or when presented with specific researcher-specified tasks.
In educational contexts, educators, teacher trainers and evaluators use video to record what other teachers do to help improve teaching practices or evaluate performance.
Researchers may record videos of learners in classrooms or museums to understand what teachers and docents do and how audiences respond.
In lab, classroom, or museum settings, video-based eye tracking systems provide information about where people look and for how long.
While not strictly video recordings, image streams from computer-based tasks or learning-rich video games may be combined with information about user keyboard or controller inputs to study real-time perfomance.
When combined with other quantitative and qualitative measures about learning outcomes, video can provide information about what kinds of teaching techniques boost learning for what kinds of students.
Video may even be used to study engagement by students enrolled in online/web-based courses or to monitor compliance with academic integrity rules.
Because video closely mimics the multisensory experiences of live human observers, recordings collected by one person for a particular purpose may be readily understood by another person and reused for another purpose.
In other words, video requires less metadata and less manual curation than other data types, but offers greater potential for reuse.
Moreover, the success, growth, and popularity of YouTube and other video-based social media (Vimeo, Periscope, Meerkat, SnapChat) demonstrates that video garners high levels of user interest and attention for interpersonal communication and entertainment uses.
The success of these services also suggests that web-based video storage and streaming systems are now sufficiently well developed to satisfy large-scale demand.
The question for researchers and policymakers is how to seize the moment in order to realize video's substantial potential to improve teaching and learning.

\section{The Challenge of Video}

Sharing research video poses challenges.
These include technical infrastructure, research ethics, privacy, and security, and data management. 
Further, making open data sharing in education research an established norm requires changing understanding about how identifiable data might be shared and building a community of researchers and educators committed to making video data sharing work.

\subsection{Technical Challenges}

@@

Videos are large.
One hour of HD video (1080p, 1440x1080 @ 60 Hz) can consume more than 10 GB.
The size of the file varies by format and resolution, and recordings vary widely, making it essential for video storage systems to convert files to a standard format.
Many settings, like classrooms, require multiple camera views to capture the desired behaviors.
Some questions involve the collection of multiple, temporally dense data streams (physiological recordings, motion tracking, eye tracking, verbal transcripts, keyboard/controller responses).
These require tools for synchronization across camera views and data streams.
The use of video creates a data explosion: A typical lab studying infant or child development collects 8-12 hours of video/week [cite18] in widely varied formats. 
Thus, sharing digital video requires substantial storage capacity, significant computational resources for transcoding videos into common, preservable formats, and powerful search and streaming tools,.

After videos are recorded, human coders evaluate videos and apply text-based annotations or numeric codes to videos.
Coders use a wide-range of commercial (Noldus Observer, Mangold Interact, MaxQDA, V-Note), and academic (Transana, Elan, Datavyu) software tools, among others.
Some researchers use spreadsheets or pencil and paper.
Some tools (Datavyu, Mangold, V-Note) offer text-based export to interoperable file formats, but the tools have no standard format for data storage or export.
Some tools offer web/cloud based storage (e.g., Datavyu via Databrary, Transana, V-Note), but otherwise coded data are easily shared within tools, much less between them.

With few exceptions (Chen Yu; Gilmore \& Raudies, Rehg?, others), machine learning has not been applied to video data coding in the developmental and learning sciences.
But, there is explosive growth in the use of machine learning and computer vision tools to tag images and videos in commercial settings.
This may create opportunities for future projects that supply machine-generated tags to videos collected in educational and developmental research.

\subsection{Ethics and Privacy}

Sharing video recordings poses a unique challenge to existing data sharing policies because videos contain personally identifying information—specifically participants’ faces and voices and often the
insides of their homes and classrooms. 
Sharing personally identifiable information puts research participants and others depicted in recordings at increased risk for loss of privacy. 
At the same time, blurring or altering original recordings to hide identities undermines or eliminates
their value to other researchers. 
Often, participants’ faces and voices produce the behaviors of interest.
A challenge for future education research is to find ways to preserve the value of video for reuse by others while protecting the privacy of research participants.
Databrary has developed an access model described below that provides an example of how this might be accomplished.
\subsection{Data Management}
A primary impediment to sharing is researchers’ lack of time to find, label, clean, organize, and copy their files into formats that can be used and understood by others15. 
Even scientists committed to data sharing find it difficult to do. Data management systems that make sharing convenient, reliable—and optimally automatic29—can help.

Existing data management practices pose challenges for video sharing.
Despite similar research methods, study designs vary widely. 
No two developmental science labs manage data in the same way. 
Some studies are longitudinal, where researchers observe the same participants at
multiple sessions. 
Some are cross-sectional, where researchers observe
each participant at only one session. 
The timing of observations may be
determined by participants’ age (e.g., 4-month-olds, 8-month-olds,
12-month-olds) or abilities (e.g., preverbal, one-word utterances,
sentences), experimentally determined variables (e.g., pre- and
post-intervention), or other factors. 
Some labs organize longitudinal data by grouping files first by participant and then by session date.
Other labs organize longitudinal data by session and then by
participant. 
Still others organize longitudinal data based on task (book reading, block building, free play). Some researchers institute a central data management system to be followed by all the lab members, providing easier access and greater transparency for the entire lab but not necessarily providing the structure for similar benefits to
researchers unfamiliar with the lab’s practices. 
Other researchers allow their students to keep separate records, making it difficult to share
data even within the lab. 
Some researchers keep videos, metadata, and analyses together, and some do not. Idiosyncratic terms, record-keeping, and data management systems are the norm. 
Databrary must enable researchers to discover and understand each other’s materials, regardless of the original investigator’s data management system.
As in other fields22, developmental science is inundated by an explosion of digital data, most of which is inaccessible to other researchers. The data deluge stems from the advent of cheap, high-resolution digital cameras. The average developmental lab collects 12 hours of video/week23. Raw data without information about workflow and data provenance is essentially “orphaned” and cannot be understood by subsequent researchers who weren’t part of the original team24. Moreover, video files must be electronically linked with participant permissions and relevant metadata. An effective data management system for video does not currently exist. We will build one to allow researchers within labs to benefit from each other’s work, facilitate interdisciplinary collaborations, and provide a solid foundation for open sharing25

@@

\section{Databrary.org}
Databrary (databrary.org) is a digital data library specialized for storing and sharing video and research data and metadata.
Databrary is a joint project of New York University (NYU) and The Pennsylvania State University (Penn State). 
The Databrary project began with a workshop on open video data sharing funded by the National
Science Foundation (NSF) under Grant No. BCS-1139702 at which developmental scientists, computer scientists, library scientists, and federal agency program officers discussed the promises and challenges of
sharing video data. 
NSF and the National Institute of Child Health and Human Development (NICHD) have provided funding under grant No. BCS-1238599 and Cooperative Agreement U01-HD-076595, respectively. 
The interdisciplinary team bring expertise in developmental and neural
science, information technology, library science, software development,
data curation, and community engagement. 
In addition, a board of expert
advisors composed of developmental scientists, library scientists, and
leaders of data repositories in the behavioral and social sciences
provide guidance.
No large-scale repository for sharing video data currently exists. So, we will create a web-based Databrary repository.

@@

Databary is built on a PostreSQL database using the Scala Play framework and JavaScript.
Data are preserved indefinitely in a secure data storage facility at NYU managed by central IT.
There is no cost to use the system; an institutional subscription model is under development.
All code is hosted on GitHub.
Databrary can house video, audio, PDF, spreadsheet, image, and text-based files (along with associated metadata), as well as executable scripts in Ruby, R and Matlab that facilitate data analysis.
Video and audio data are transcoded into standard and HTML5-compatible formats, currently H.264+AAC as MP4. Databrary stores other data in native formats (e.g., .doc, .docx, .xls, .xlsx, .txt, .csv, .pdf, .jpg, .png).
